---
title: "Which Democratic candidates are getting mentioned the most on cable and online news?"
author: "Tanner Gildea"
date: "7/17/2019"
output: html_document
---

<br />

FiveThirtyEight is currently running an [interesting series](https://fivethirtyeight.com/tag/media/) analyzing how cable news and online media sources are covering the Democratic presidential candidates. [Dhrumil Mehta](https://fivethirtyeight.com/contributors/dhrumil-mehta/), one of FiveThirtyEight's data journalists, has  diligently updated the series each week, keeping track of how coverage of individual candidates has risen and fallen. They have also kindly shared the underlying data [here](https://github.com/fivethirtyeight/data/tree/master/media-mentions-2020).

Personally, I was interested in seeing how coverage of the top candidates has changed since the beginning of the race (roughly January of 2019). I also wanted to see if coverage of candidates diverged based on media type - cable or online. 

The graph below helps answer these two questions.

<br />

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 8)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(janitor)
library(readxl)
library(gt)
library(ggthemes)
library(data.table)


download.file(url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/media-mentions-2020/cable_weekly.csv",
              destfile  = "~/Desktop/R.projects/dems_media_coverage/cable_weekly.csv",
              mode = "wb")    

download.file(url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/media-mentions-2020/online_weekly.csv",
              destfile  = "~/Desktop/R.projects/dems_media_coverage/online_weekly.csv",
              mode = "wb")   

cable <- read_csv("cable_weekly.csv")


online <- read_csv("online_weekly.csv")

fs::file_delete(c("cable_weekly.csv", "online_weekly.csv"))



cable_joined_online <- merge(cable, online, by = c("name","date" ))
```

```{r join}
clean_cable_joined_online <- cable_joined_online %>% 
  filter(name %in% c("Joe Biden", 
                     "Pete Buttigieg", 
                     "Elizabeth Warren", 
                     "Kamala Harris", 
                     "Bernie Sanders",
                     "Beto O'Rourke",
                     "Julian Castro",
                     "Amy Klobuchar",
                     "Cory Booker")) %>% 
  group_by(name)

# Needed to distinguish on graph w/ fill and color

clean_cable_joined_online$query.x <- "Cable News"
clean_cable_joined_online$query.y <- "Online Articles"

# Start graph

ggplot(clean_cable_joined_online, aes(x = date, 
                        y = pct_of_all_candidate_stories,
                        color = query.x,
                        fill = query.x)) +
  
  geom_point() +
  
  geom_smooth(level  = .50, method = "loess") +
  
  # for second line
  
  geom_point(aes(x = date, 
                        y = pct_of_all_candidate_clips,
                        color = query.y,
                        fill = query.y)) +
  
  geom_smooth(aes(x = date, 
                        y = pct_of_all_candidate_clips,
                        color = query.y,
                        fill = query.y),
              
              level  = .50, method = "loess") +
  
  # Makes scale more readable
  
  scale_y_sqrt(labels = scales::percent,
    breaks = c(.01, .025, .05, 0.1, .3, 0.2, 0.4, .5, .6)) +
  
  facet_wrap(~name) +
  
#  scale_x_date(date_breaks = "1 month") + 
  
  # Style Settings
  
  labs(x = NULL, y = "Percent of Articles and Stories That Mention Each Candidate",
       title = "How often are the candidates mentioned in cable and online news?",
       subtitle = "Cable includes CNN, Fox News, and MSNBC - Online articles sourced from the Media Cloud database",
       caption = "Source: FiveThirtyEight.com") +
  
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  
  scale_fill_discrete(name = "Media Type") + 
  scale_color_discrete(name = "Media Type") +
  
  
  theme_tufte() +
  
  theme(plot.title = element_text(size = rel(1.5), face ="bold"),
        plot.subtitle = element_text(face = "italic"),
        legend.title = element_text(size = rel(1.2), face = "bold"),
        axis.title.y = element_text(size = rel(1.2)),
        plot.caption = element_text(face = "italic"),
        panel.grid.major.y = element_line(color = "gray"),
        axis.line.x = element_line(), axis.line.y = element_line())

```


```{r change_table}

# Cable table

cable_table <- cable %>% 
  select(name, date, pct_of_all_candidate_clips) %>% 
  filter(date == "2019-01-06" | date == "2019-07-07") %>% 
  spread(key = date, value = pct_of_all_candidate_clips) %>% 
  setnames(old=c("2019-01-06","2019-07-07"), new=c("Cable 1", "Cable 2")) 

cable_table <- cable_table %>% 
  mutate(change = cable_table$'Cable 2' - cable_table$'Cable 1')

# Online table

online_table <- online %>% 
  select(name, date, pct_of_all_candidate_stories) %>% 
  filter(date == "2019-01-06" | date == "2019-07-07") %>% 
  spread(key = date, value = pct_of_all_candidate_stories) %>% 
  setnames(old=c("2019-01-06","2019-07-07"), new=c("Online 1", "Online 2"))

as.numeric(as.character(cable_table$'Online 1'))
as.numeric(as.character(cable_table$'Online 2'))

joined_table <- left_join(cable_table, online_table) %>% 
  mutate()

```






```{r cable_graph}

# clean_cable <- cable %>% 
#   filter(name %in% c("Joe Biden", 
#                      "Pete Buttigieg", 
#                      "Elizabeth Warren", 
#                      "Kamala Harris", 
#                      "Bernie Sanders",
#                      "Beto O'Rourke")) %>% 
#   group_by(name)
# 
# 
# ggplot(clean_cable, aes(x = date, 
#                         y = pct_of_all_candidate_clips,
#                         color = name,
#                         fill = name)) +
#   
#   geom_point() +
#   geom_smooth(level  = .50, method = "loess") +
#   
#   scale_y_sqrt(labels = scales::percent,
#     breaks = c(.01, .025, .05, 0.1, 0.2, 0.4, .6)) +
#    
#   facet_wrap(~name) +
#   
# #  scale_x_date(date_breaks = "1 month") + 
#   
#   labs(x = NULL, y = "Percent of Coverage",
#        title = "Cable Attention Paid to Top Democratic Candidates",
#        subtitle = "Averages Greater than 5% Across CNN, Fox News, and MSNBC",
#        caption = "Source: FiveThirtyEight.com") +
#   
#   scale_color_brewer(palette = "Paired") +
#   scale_fill_brewer(palette = "Paired") +
#   
#   scale_fill_discrete(name = "Candidate") + 
#   scale_color_discrete(name = "Candidate") +
#   
#   theme_tufte() +
#   
#   theme(plot.title = element_text(size = rel(1.5), face ="bold"),
#         plot.subtitle = element_text(face = "italic"),
#         legend.title = element_text(size = rel(1.2), face = "bold"),
#         axis.title.y = element_text(size = rel(1.2)),
#         plot.caption = element_text(face = "italic"),
#         panel.grid.major.y = element_line(color = "gray"))
```

```{r online_graph}
# clean_online <- online %>% 
#   filter(name %in% c("Joe Biden", 
#                      "Pete Buttigieg", 
#                      "Elizabeth Warren", 
#                      "Kamala Harris", 
#                      "Bernie Sanders",
#                      "Cory Booker")) %>% 
#   group_by(name)
# 
# 
# ggplot(clean_online, aes(x = date, 
#                         y = pct_of_all_candidate_stories,
#                         color = name,
#                         fill = name)) +
#   
#   geom_point() +
#   geom_smooth(level  = .50, method = "loess") +
#   
#   scale_y_sqrt(labels = scales::percent,
#     breaks = c(.025, .05, 0.1, 0.2, 0.4, .6)) +
#   
#   facet_wrap(~name) +
#   
# #  scale_x_date(date_breaks = "1 month") + 
#   
#   labs(x = NULL, y = "Percent of Coverage",
#        title = "Online Attention Paid to Top Democratic Candidates",
#        subtitle = "Averages Greater Than 10% in the Media Cloud database",
#        caption = "Source: FiveThirtyEight.com") +
#   
#   scale_color_brewer(palette = "Set1") +
#   scale_fill_brewer(palette = "Set1") +
#   
#   scale_fill_discrete(name = "Candidate") + 
#   scale_color_discrete(name = "Candidate") +
#   
#   theme_tufte() +
#   
#   theme(plot.title = element_text(size = rel(1.5), face ="bold"),
#         plot.subtitle = element_text(face = "italic"),
#         legend.title = element_text(size = rel(1.2), face = "bold"),
#         axis.title.y = element_text(size = rel(1.2)),
#         plot.caption = element_text(face = "italic"),
#         panel.grid.major.y = element_line(color = "gray"))


```


```{r yang}
# clean_yang <- online %>% 
#   filter(name %in% c("Andrew Yang", "Marianne Williamson", "Eric Swalwell", "Tulsi Gabbard")) %>% 
#   group_by(name)
# 
# 
# ggplot(clean_yang, aes(x = date, 
#                         y = pct_of_all_candidate_stories,
#                         color = name,
#                         fill = name)) +
#   
#   geom_point() +
#   geom_smooth(level  = .50, method = "loess") +
#   
#   scale_y_sqrt(labels = scales::percent,
#     breaks = c(.01, .025, .05, 0.1, 0.2, 0.4, .6)) +
#   
# #  scale_x_date(date_breaks = "1 month") + 
#   
#   labs(x = NULL, y = "Percent of Coverage",
#        title = "Online Attention Paid to Top Democratic Candidates",
#        subtitle = "From the Media Cloud database",
#        caption = "Source: FiveThirtyEight.com") +
#   
#   scale_color_brewer(palette = "Paired") +
#   scale_fill_brewer(palette = "Paired") +
#   
#   scale_fill_discrete(name = "Candidate") + 
#   scale_color_discrete(name = "Candidate") +
#   
#   theme_tufte() +
#   
#   theme(plot.title = element_text(size = rel(1.5), face ="bold"),
#         plot.subtitle = element_text(face = "italic"),
#         legend.title = element_text(size = rel(1.2), face = "bold"),
#         axis.title.y = element_text(size = rel(1.2)),
#         plot.caption = element_text(face = "italic"),
#         panel.grid.major.y = element_line(color = "gray"))
```




